The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-03-25 VegOzone.tex
Added tables for data sources and some text.

\section{Data}
\label{sec:data}
TEXT
\begin{figure}[t]
  \includegraphics[width=8.3cm]{station_map_fennoscandia}
  \caption{Cap of the North. Location of past and present ozone observation sites in northern Fennoscandia used within this study.}
  \label{fig:station_map_fennoscandia}
\end{figure}

\subsection{NIBIO research station Svanhovd}
\label{subsec:svanhovd}
"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north. It is where the eastern Siberian taiga meets the western Boreal forest. [...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\
\emph{Comment: Invite co-authors from NILU and NIBIO?}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{svanhovd_researchstation}
  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography (c) Norges Kartverk}
  \label{fig:svanhovd_research_station}
\end{figure}

\subsubsection{Ozone monitoring}
\label{subsubsec}
{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical periode. Project related monitoring (summer seasons 2018/19) $\rightarrow$ NILU -- Sverre?}

@@

\subsection{Other observational data}
\label{subsec:obs_data}

\subsubsection{EBAS database}
\label{subsubsec:ebas}

\begin{table}[t]
  \caption{Ozone observations in northern Fennoscandia available from EBAS.}
  \label{tab:ebas_obs}
\begin{tabular}{lllllrc}
\tophline
Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
           &         &         & lat             & lon               & alt            &\\
           &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
\middlehline
Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996 $^\dagger$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Exclusive data taking summer seasons 2018/19.} % Table Footnotes
\end{table}

\begin{figure*}[t]
  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
  \label{fig:ozone_timesseries_fenoscandic_obs}
  \caption{Time series of ozone observations in northern Fennoscandia. Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19. The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
\end{figure*}

\subsection{Model and reanalysis data}
Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea
Precipitation/Temperature: CRU bias corrected rea
Ozone uptake / POD: DO3SE

\section{Statistical analysis}
TEXT

>>>>

\section{Data}
\label{sec:data}

In our assessment on ozone damage on vegetation is the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley. For the purpose of this study, we evaluate data taken at Svanhovd with observational data (Section~\ref{subsec:svanhovd}) with different monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}). Reanalysis data is used for complementation were observational data was not available or temporally sparse (Section~\ref{subsec:model_data}).

\subsection{NIBIO research station Svanhovd}
\label{subsec:svanhovd}
"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north. It is where the eastern Siberian taiga meets the western Boreal forest. [...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\ The research station and locations of the atmospheric monitoring site and the ozone garden are shown in the aerial photography (Fig.~\ref{fig:station_map_fennoscandia}). Svanhovd comprises atmospheric monitoring (ozone monitoring exclusively for the summer seasons 2018/19) and observed ozone damage in plants.
\emph{Comment: Invite co-authors from NILU and NIBIO?}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{svanhovd_researchstation}
  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
  \label{fig:svanhovd_research_station}
\end{figure}

\subsubsection{Ozone monitoring}
\label{subsubsec}
{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical periode. Project related monitoring (summer seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}

@@

\subsection{Other observational data}
\label{subsec:obs_data}
For the purpose of this study, we are interested in past and present observations of surface ozone concentrations in northern Fennoscandia. This limits the available observation data to a handfull of atmospheric monitoring sites with data available from the NILU EBAS database. An overview over the geographical locations is given in Fig.~\ref{fig:station_map_fennoscandia}.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{station_map_fennoscandia}
  \caption{Cap of the North. Location of past and present ozone observation sites in northern Fennoscandia used within this study.}
  \label{fig:station_map_fennoscandia}
\end{figure}

\subsubsection{Ozone monitoring}
\label{subsubsec:ebas}

\begin{table}[t]
  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on EBAS.}
  \label{tab:ebas_obs}
\begin{tabular}{lllllrl}
\tophline
Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
           &         &         & lat             & lon               & alt            &\\
           &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
\middlehline
Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Exclusive ozone monitoring summer seasons 2018/19 operated by NILU.} % Table Footnotes
\end{table}

\begin{figure*}[t]
  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
  \label{fig:ozone_timesseries_fenoscandic_obs}
  \caption{Time series of ozone observations in northern Fennoscandia. Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19 (Tab.~\ref{tab_ebas_obs}). The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
\end{figure*}

\subsubsection{Atmospheric monitoring}
Monitoring data from Svanvik/Pasvik.
Ozone/Temperature (2018/2019): Luftkvalitet.no (NILU), LandbruksMeteorologiskTjeneste (NIBIO)

\subsection{Reanalysis data}
\label{subsec:model_data}
Range of reanlysis data.
\subsubsection{Ozone}

\begin{table}[t]
\caption{Ozone related ranalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
\begin{tabular}{llccl}
\tophline
Name & Organization & \multicolumn{2}{c}{Resolution} & Time period \\
&              & spatial & temporal & \\
\middlehline
MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012\\
CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012$^\dagger$\\
CAMS Regional Air Quality & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018$^\dagger$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Subset of the data used in this study. The full dataset extends into near real time.} % Table Footnotes
\end{table}

Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea ensemble mean

\subsubsection{Temperature and precipitation}
Precipitation/Temperature: CRU bias corrected rea

\section{Statistical analysis}
Statistical evaluation of Svanhovd (Svanvik/Pasvik) ozone data in a regional and climatological context.
